Hydra Battles and Cie

### 2020-11-15 doc/chapter-powers.tex
Added addition-chains stuff into a general hydras-and-Cie development
- Not tested yet :
  - opam directives, etc.
     the current states uses coq-equations, coq-param, and
         coq-mathcomp-algebra (for the contribution by Yves)

  - There is an ml file, provisionnaly stored as theories/additions/stub.ml,
    and a shell script: theories/additions/make_fib_tests.txt
    stub.ml uses two files generated by extraction from fib.v : bigfib.ml and bigfig.mli.

\emph{From Module 
\href{../src/html/addition.power.FirstSteps.html}{\texttt{Powers.FirstSteps}}}
\label{sect: power-definitions}

@@

\emph{From Module
\href{../src/html/addition.power.FirstSteps.html}{Powers.FirstSteps}}

@@

@@

\emph{From Module \href{../src/html/addition.power.Monoid_def.html}{Powers/Monoid\_def.v}.}

@@

\emph{From Module \href{../src/html/addition.power.Monoid_def.html}{Powers.Monoid\_def}}

@@

The following class definition, from Module \href{../src/html/addition.power.Monoid_def.html}{Powers.Monoid\_def},
is parameterized with some type $A$,
a multiplication (called \texttt{op} in the definition), and a neutral element
$\mathds{1}$ (called \texttt{one} in the definition).

@@

Let us show various instances, which will be used in further proofs and examples.
Complete definitions and proofs are given in 
File~\href{../src/html/addition.power.Monoid_instances.html}{Powers/Monoid\_instances.v}.

@@

\noindent
\emph{From Module \href{../src/html/addition.power.Monoid_def.html}{Powers.Monoid\_def}}

@@

\emph{From Module \href{../src/html/addition.power.Monoid_instances.html}{Powers.Monoid\_instances}}

@@

The  module \href{../src/html/addition.power.Pow.html}{Powers.Pow} defines two functions for exponentiation on any 
\texttt{EMonoid}  on carrier $A$.
They are essentially the same as in Sect.~\vref{sect: power-definitions}. The main difference lies in the arguments of the functions, which now contain
 an instance~\texttt{M} of class \texttt{EMonoid}. 
Thus, the arguments associated with the multiplication,
the neutral element and the equivalence relation associated with \texttt{M}
are left implicit.

@@

From Module~\href{../src/html/addition.power.Demo_power.html}{Powers.Demo\_power}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/addition.power.Pow.html}{Powers.Pow}}

@@

\vspace{4pt}
From Module~\href{../src/html/addition.power.Pow.html}{Powers.Pow}

@@

\subsection{Proof of correctness of binary exponentiation w.r.t. the function \texttt{power}}
Section \texttt{M\_given} of Module 
~\href{../src/html/addition.power.Pow.html}{Powers.Pow} is devoted to the proof 
of properties of the functions above.
Note that properties of \texttt{power} refer to the \emph{specification} of exponentiation, and can be applied for proving correctness of any implementation.

@@

From Module~\href{../src/html/addition.power.Chains.html}{Powers.Chains}

@@

This tactic is not adapted to much bigger exponents. In \linebreak
 Module~\href{../src/html/addition.power.Euclidean_Chains.html}{Euclidean\_Chains},
 for instance, we tried to apply this tactic for proving the correctness 
of a chain associated with the exponent $45319$. 
 We had to interrupt the prover, which 
was trying to build a linear tree of $2\times  45319 + 1$ nodes!
Indeed, using \texttt{reflection\_correct\_tac} is like doing a 
symbolic evaluation of an inefficient (linear) exponentiation algorithm.

@@

\paragraph{Note:}
All the \coq{} material described in this section is available on 
 Module~\href{../src/html/addition.power.Euclidean_Chains.html}{Powers/Euclidean\_Chains.v}

@@

Module ~\href{../src/html/addition.power.Dichotomy.html}{Powers.Dichotomy} contains a definition of the function \texttt{delta}, and proofs that if 
$n>3$ then $1<\delta(n)<n$.

>>>>

\emph{From Module 
\href{../theories/html/hydras.additions.FirstSteps.html}{\texttt{additions.FirstSteps}}}
\label{sect: power-definitions}

@@

\emph{From Module
\href{../theories/html/hydras.additions.FirstSteps.html}{additions.FirstSteps}}

@@

\begin{todo}
Document the files contributed by Yves
\begin{itemize}
\item additions/fib.v (to rename ?)
\item additions/stub.ml (to keep inside theories/ or move to src/ ?)
\item theories/additions/make\_fib\_tests.txt (to put in a Makefile?)
\end{itemize}
\end{todo}

@@

\emph{From Module \href{../theories/html/hydras.additions.Monoid_def.html}{additions/Monoid\_def.v}.}

@@

\emph{From Module \href{../theories/html/hydras.additions.Monoid_def.html}{additions.Monoid\_def}}

@@

The following class definition, from Module \href{../theories/html/hydras.additions.Monoid_def.html}{additions.Monoid\_def},
is parameterized with some type $A$,
a multiplication (called \texttt{op} in the definition), and a neutral element
$\mathds{1}$ (called \texttt{one} in the definition).

@@

Let us show various instances, which will be used in further proofs and examples.
Complete definitions and proofs are given in 
File~\href{../theories/html/hydras.additions.Monoid_instances.html}{additions/Monoid\_instances.v}.

@@

\noindent
\emph{From Module \href{../theories/html/hydras.additions.Monoid_def.html}{additions.Monoid\_def}}

@@

\emph{From Module \href{../theories/html/hydras.additions.Monoid_instances.html}{additions.Monoid\_instances}}

@@

The  module \href{../theories/html/hydras.additions.Pow.html}{additions.Pow} defines two functions for exponentiation on any 
\texttt{EMonoid}  on carrier $A$.
They are essentially the same as in Sect.~\vref{sect: power-definitions}. The main difference lies in the arguments of the functions, which now contain
 an instance~\texttt{M} of class \texttt{EMonoid}. 
Thus, the arguments associated with the multiplication,
the neutral element and the equivalence relation associated with \texttt{M}
are left implicit.

@@

From Module~\href{../theories/html/hydras.additions.Demo_power.html}{additions.Demo\_power}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.additions.Pow.html}{additions.Pow}}

@@

\vspace{4pt}
From Module~\href{../theories/html/hydras.additions.Pow.html}{additions.Pow}

@@

\subsection{Proof of correctness of binary exponentiation w.r.t. the function \texttt{power}}
Section \texttt{M\_given} of Module 
~\href{../theories/html/hydras.additions.Pow.html}{additions.Pow} is devoted to the proof 
of properties of the functions above.
Note that properties of \texttt{power} refer to the \emph{specification} of exponentiation, and can be applied for proving correctness of any implementation.

@@

From Module~\href{../theories/html/hydras.additions.Chains.html}{additions.Chains}

@@

This tactic is not adapted to much bigger exponents. In \linebreak
 Module~\href{../theories/html/hydras.additions.Euclidean_Chains.html}{Euclidean\_Chains},
 for instance, we tried to apply this tactic for proving the correctness 
of a chain associated with the exponent $45319$. 
 We had to interrupt the prover, which 
was trying to build a linear tree of $2\times  45319 + 1$ nodes!
Indeed, using \texttt{reflection\_correct\_tac} is like doing a 
symbolic evaluation of an inefficient (linear) exponentiation algorithm.

@@

\paragraph{Note:}
All the \coq{} material described in this section is available on 
 Module~\href{../theories/html/hydras.additions.Euclidean_Chains.html}{additions/Euclidean\_Chains.v}

@@

Module ~\href{../theories/html/hydras.additions.Dichotomy.html}{additions.Dichotomy} contains a definition of the function \texttt{delta}, and proofs that if 
$n>3$ then $1<\delta(n)<n$.
